Proofs of Linearizability Based on Forward Simulations

### 2017-01-23 draftCAV/prelim.tex
dealing with simulations

%\textcolor{red}{For the below paragraph, I cannot see the equivalence  of informal explanation and formal definition of weakening. I think the formal one should be if a ret comes before a call in $h_1'$ then this order is preserved in $h_2$. But the definition says iff. Ignore this if I am wrong.}
The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.
Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. This weakening relation
between histories can be formalized as follows: a history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
if{f} there exists a well-formed history $h_1'$
obtained from $h_1$ by appending return actions, and deleting call actions,
s.t.:
\begin{quote}

  $h_2$ is a permutation of $h_1'$ that preserves the order between
  return and call actions, i.e.,~if a given return action occurs before a given
  call action in $h_1'$, then the same holds in $h_2$.

\end{quote}
%A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).

@@

%\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events. Then, we put forward the idea of enriching $\Gamma$ by adding new actions. In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained. Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked. I propose to first introduce linearizability. Then explain refinement and give Theorem 1.}
Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
is formally captured by \emph{(observational) refinement}. Informally, given two libraries
$L_1$ and $L_2$ implementing the methods of some concurrent object, we
say $L_1$ refines $L_2$ if{f} every computation of every program
using $L_1$ would also be possible were $L_2$ used instead. We assume that a program can 
interact with the library only through call and return actions, and thus refinement can be defined
as history set inclusion. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
%We fix a set of methods $\<Methods>$ and values $\<Vals>$.

>>>>

%\textcolor{red}{For the below paragraph, I cannot see the equivalence  of informal explanation and formal definition of weakening. I think the formal one should be if a ret comes before a call in $h_1'$ then this order is preserved in $h_2$. But the definition says iff. Ignore this if I am wrong.}
The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.
Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. A 
history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
if{f} there exists a history $h_1'$
obtained from $h_1$ by appending return actions, and deleting call actions,
s.t.:
  $h_2$ is a permutation of $h_1'$ that preserves the order between
  return and call actions, i.e.,~if a given return action occurs before a given
  call action in $h_1'$, then the same holds in $h_2$.
%A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).

@@

%\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events. Then, we put forward the idea of enriching $\Gamma$ by adding new actions. In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained. Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked. I propose to first introduce linearizability. Then explain refinement and give Theorem 1.}
Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
is formally captured by \emph{(observational) refinement}. Informally, given two libraries
$L_1$ and $L_2$ implementing the methods of some concurrent object, we
say $L_1$ refines $L_2$ if{f} every computation of every program
using $L_1$ would also be possible were $L_2$ used instead. We assume that a program can 
interact with the library only through call and return actions, and thus refinement can be defined
as history set inclusion. Refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} 
when $L_2$ is an atomic library~\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
. 
%We fix a set of methods $\<Methods>$ and values $\<Vals>$.

### 2017-01-23 draftCAV/refProofs.tex
dealing with simulations

\begin{dfn}
Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ 
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $\delta_2(u,\gamma, u')$. 
\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s|$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation shows that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$. In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action. The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.

The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations. It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.
\begin{theorem}\label{th:forSim}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. If there is a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$. Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
\end{theorem}

>>>>

\begin{dfn}
Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item $\mathit{fs}[s_0^1] = \{s_0^2 \}$.
\item If $(s,\gamma,s') \in \delta_1$, for some $\gamma\in \Gamma$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for each $j\neq i$.
\item If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation requires that every step of $L_1$ corresponds to a sequence of steps of $L_2$. To imply $\Gamma$-refinement, every step of $L_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $L_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.

The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations (when $L_2$ is $\Gamma$-deterministic). It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.
\begin{theorem}\label{th:forSim}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. If there is a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$. Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
\end{theorem}
